Journal and article coverage

### 2025-02-19 analysis/manuscript.qmd
Check missing dois

hoaddata, developed and maintained by the author to support open access monitoring and research [@Jahn_2025], is an R data package comprising information about the uptake of hybrid open access since 2017 collected from several openly available data sources.
It combines article-level metadata from Crossref and OpenAlex with transformative agreement information from the cOAlition S Journal Checker Tool (JCT), which links journal and institutional data to agreements in the ESAC registry.

@@

<!-- The JCT journal data was enriched with the linking ISSN (ISSN-L) according to the ISSN Registry Agency.
A comprehensive exclusion of fully open access journals was performed, following the methodology described in @jahn_2025.
This involved checking multiple journal lists (DOAJ, OpenAlex, Bielefeld) and article-level investigations.
The JCT institution data was enriched with ROR-IDs from associated institutions, such as university hospitals or institutes of large research organisations such as the Max Planck Society, according to OpenAlex' institution entity.
<!-- In total, xxxx hybrid journals could be linked to xxx agreements listed in the ESAC Transformative Agreements Registry, which formed to basis for obtaining article-level data including author affiliations, open access status ans document types. -->
Article metadata was retrieved from each databse using all ISSN variants linked to an ISSN-L. The metadata included DOIs, publication dates, article types, open access information as well as author roles and institutional affiliations.
Publication years were determined using the earliest known date of publication in a journal.
articles in hoaddata, this corresponded to Crossref's issued date.<!--, although variations were observed (check ref)-->
For Web of Science and Scopus, the earliest publication date was used where available, with Scopus dates specifically determined by the KB through version tracking of the raw data.

@@

While hoaddata only covers articles with a DOI, Scopus and Web of Science publication metrics were aggregated using the database identifier.
A subsequent comparison of DOI coverage shows that non-core articles in Web of Science often lacked a DOI.
This was particularly the case for meeting abstracts, which are notably prevalent in Health Sciences journals [@Melero_Fuentes_2025].
Open access indicators were aggregated by DOI, as Unpaywall only collects information on open access status for articles with a DOI.
A closer look at the core articles with affiliation data reveals a lack of data on corresponding authors in the case of OpenAlex, the affiliation data source used by hoaddata, compared to Web of Science and Scopus.
Only about two-thirds of the articles examined provided corresponding author affiliations.
For first authors, the proportion was `r methods_overview_table_df |> filter(indicator == "First author articles") |> pull(hoad) / methods_overview_table_df |> filter(indicator == "Core articles with DOI") |> pull(hoad) * 100`%.
Therefore, only first author data for hoaddata were considered in the following analysis.

@@

upset_overview <- ComplexUpset::upset(
  upset_data,
  c("Crossref", "WoS", "Scopus"),
  # Only journals in Crossref
  min_size = 30,
  set_sizes = FALSE,
  stripes = "transparent",
  # wrap = TRUE,
  name = "Data Source Coverage",
  base_annotations = list(
    "Journals" = (
      ComplexUpset::intersection_size(
        counts = FALSE,
        text = list(size = 8 / .pt),
        text_mapping = aes(
          label = paste0(!!upset_text_percentage(), '\n(', format(!!size, big.mark = ","), ')'),
          colour = ifelse(!!size > 5000, 'on_bar', 'on_background'),
          y = ifelse(!!size > 5000, !!size - 4000, !!size)
        ), width = bar_width
      ) +
        scale_y_continuous(labels = scales::number_format(big.mark = ",")) +
        ggtitle("A")
    )
  ),
  annotations = list(
    "Journal size\n(log scale)" = (
      ggplot(mapping = aes(x = intersection, y = n)) +
        geom_boxplot(outliers = FALSE, width = bar_width) +
        scale_y_log10(labels =  scales::number_format(big.mark = ",")) +
        ggtitle("C")
    ),
    "Articles" = (
      ggplot(mapping = aes(y = n_cr)) +
        geom_bar(
          stat = "identity",
          color = "grey80",
          width = bar_width,
          aes(fill = "All")
        ) +
        geom_bar(
          aes(y = overlap, fill = "Shared"),
          stat = "identity",
          width = bar_width,
          color = "#0091c8"
        ) +
        scale_fill_manual(
          values = c(
            `All` = "grey80",
            `Shared` = "#0091c8"
          ),
          name = "DOI"
        ) +
        scale_y_continuous(labels =  scales::number_format(big.mark = ","),
                           breaks = c(0, 2500000, 5000000, 7500000)) +
        ggtitle("B")
    )
  ),
  matrix = (intersection_matrix(geom = geom_point(
    shape = "square", size = 2
  )))
)
upset_overview & my_upset_theme
```
Journal coverage analysis revealed that `r round(nrow(union_jns) / nrow(all_jns) * 100)`% (n = `r nrow(union_jns)`) of hybrid journals included in transformative agreements were indexed in all three databases ( \ref{fig-upset_coverage_results}A).
The second-largest set consisted of journals indexed in both hoaddata and Scopus, comprising `r round( jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == TRUE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == TRUE) |>  nrow()`) of hybrid journals.
Notably, `r round(jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |>  nrow()`) of journals were exclusively contained in hoaadata, while `r round( jn_upset_articles |> filter(WoS == TRUE, Crossref == FALSE, Scopus == TRUE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(WoS == TRUE, Crossref == FALSE, Scopus == TRUE) |>  nrow()`) were only found in the proprietory databases Scopus and Web of Science.
Upon insepction, this latter group mainly represents hybrid journals for which no open access evidence could be retrieved from Crossref, which serves as the open access evidence source for hoaddata.

In terms of article coverage, Fig B shows the DOI overlap, distinguihsing between a shared corpus and just those where a DOI was only available through hoaddata.
Upon inspection, reasons for lackign DOI coverage in proprietory database were different publication dates and differenc ein document type definitions. 
As such, demosntrates certain uncertainies using the open approach.
Overall, The majority of articles can be found

@@

This study compares results drawn from the openly available and regularly updated hoaddata, an R package leveraging the open scholarly data sources Crossref and OpenAlex, with the in-house database of the German Competence Network 
Overall, xxxx hybrid journals included in xxx agreements that published at least one open access article between 2019 and 2023.

>>>>

hoaddata, developed and maintained by the author to support open access monitoring and research [@Jahn_2025], is an R data package collecting information on hybrid open access uptake from multiple openly available data sources.
It combines article-level metadata from Crossref and OpenAlex with transformative agreement information from the cOAlition S Journal Checker Tool (JCT), which links journal and institutional data to agreements in the ESAC registry.

@@

<!-- The JCT journal data was enriched with the linking ISSN (ISSN-L) according to the ISSN Registry Agency.
A comprehensive exclusion of fully open access journals was performed, following the methodology described in @jahn_2025.
This involved checking multiple journal lists (DOAJ, OpenAlex, Bielefeld) and article-level investigations.
The JCT institution data was enriched with ROR-IDs from associated institutions, such as university hospitals or institutes of large research organisations such as the Max Planck Society, according to OpenAlex' institution entity.
<!-- In total, xxxx hybrid journals could be linked to xxx agreements listed in the ESAC Transformative Agreements Registry, which formed to basis for obtaining article-level data including author affiliations, open access status ans document types. -->
Article metadata were retrieved from each database using all ISSN variants linked to an ISSN-L. The metadata included DOIs, publication dates, article types, open access information as well as author roles and institutional affiliations.
Publication years were determined using the earliest known date of publication in a journal.
articles in hoaddata, this corresponded to Crossref's issued date.<!--, although variations were observed (check ref)-->
For Web of Science and Scopus, the earliest publication date was used where available, with Scopus dates specifically determined by the KB through version tracking of the raw data.

@@

While hoaddata only covers articles with a DOI, Scopus and Web of Science publication metrics were aggregated using the database identifier.
A subsequent comparison of DOI coverage shows that non-core articles in Web of Science often lacked a DOI.
This was particularly the case for meeting abstracts, which are notably prevalent in Health Sciences journals [@Melero_Fuentes_2025] and are not indexed by Scopus [@Donner_2017].
Open access indicators were aggregated by DOI, as Unpaywall only collects information on open access status for articles with a DOI.
A closer look at the core articles with affiliation data reveals a lack of data on corresponding authors in the case of OpenAlex, the affiliation data source used by hoaddata, compared to Web of Science and Scopus.
Only about two-thirds of the articles examined provided corresponding author affiliations.
For first authors, the proportion was `r methods_overview_table_df |> filter(indicator == "First author articles") |> pull(hoad) / methods_overview_table_df |> filter(indicator == "Core articles with DOI") |> pull(hoad) * 100`%.
Therefore, only first author data for hoaddata were considered in the following analysis.

@@

upset_overview <- ComplexUpset::upset(
  upset_data,
  c("Crossref", "WoS", "Scopus"),
  # Only journals in Crossref
  min_size = 30,
  set_sizes = FALSE,
  stripes = "transparent",
  # wrap = TRUE,
  name = "Data Source Coverage",
  base_annotations = list(
    "Journals" = (
      ComplexUpset::intersection_size(
        counts = FALSE,
        text = list(size = 8 / .pt),
        text_mapping = aes(
          label = paste0(!!upset_text_percentage(), '\n(', format(!!size, big.mark = ","), ')'),
          colour = ifelse(!!size > 5000, 'on_bar', 'on_background'),
          y = ifelse(!!size > 5000, !!size - 4000, !!size)
        ), width = bar_width
      ) +
        scale_y_continuous(labels = scales::number_format(big.mark = ",")) +
        ggtitle("A")
    )
  ),
  annotations = list(
    "Journal size\n(log scale)" = (
      ggplot(mapping = aes(x = intersection, y = overlap)) +
        geom_boxplot(outliers = FALSE, width = bar_width) +
        scale_y_log10(labels =  scales::number_format(big.mark = ",")) +
        ggtitle("C")
    ),
    "Articles" = (
      ggplot(mapping = aes(y = n_cr)) +
        geom_bar(
          stat = "identity",
          color = "grey80",
          width = bar_width,
          aes(fill = "Surplus")
        ) +
        geom_bar(
          aes(y = overlap, fill = "Shared"),
          stat = "identity",
          width = bar_width,
          color = "#0091c8"
        ) +
        scale_fill_manual(
          values = c(
            `Surplus` = "grey80",
            `Shared` = "#0091c8"
          ),
          name = "DOI"
        ) +
        scale_y_continuous(labels =  scales::number_format(big.mark = ","),
                           breaks = c(0, 2500000, 5000000, 7500000)) +
        ggtitle("B")
    )
  ),
  matrix = (intersection_matrix(geom = geom_point(
    shape = "square", size = 2
  )))
)
upset_overview & my_upset_theme
```
Journal coverage analysis revealed that `r round(nrow(union_jns) / nrow(all_jns) * 100)`% (n = `r nrow(union_jns)`) of hybrid journals included in transformative agreements were indexed in all three databases (Figure \ref{fig-upset_coverage_results}A).
The second-largest set consisted of journals indexed in both hoaddata and Scopus, comprising `r round( jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == TRUE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == TRUE) |>  nrow()`) of hybrid journals.
Notably, `r round(jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |>  nrow()`) of journals were exclusively contained in hoaadata, while `r round( jn_upset_articles |> filter(Crossref == FALSE, Scopus == TRUE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(Crossref == FALSE, Scopus == TRUE) |>  nrow()`) were only found in the proprietory database Scopus. 
Of these, `r jn_upset_articles |> filter(WoS == TRUE, Crossref == FALSE, Scopus == TRUE) |>  nrow()` were also available in the Web of Science.
Upon inspection, this group of hybrid journals included only in the proprietary data sources mainly represents hybrid journals for which no open access evidence could be retrieved from Crossref, which served as the open access evidence source for hoaddata.

In terms of article coverage, Figure \ref{fig-upset_coverage_results}B shows the total publication volume per combination in terms of DOI availability.
The largest set of hybrid journals, which includes all three data sources, also contains the largest number of articles. 
In total, these journals recorded `r upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum()` overlapping articles, represented by the blue bar. They represent `r round(upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum() / methods_overview_table_df |> filter(indicator == "Core articles with DOI") |> pull(wos) * 100)`% of core articles with DOI indexed in the Web of Science sample, and `r round(upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum() / methods_overview_table_df |> filter(indicator == "Core articles with DOI") |> pull(scp) * 100)`% in Scopus.
Another `r upset_data |> filter(Crossref == TRUE, WoS == FALSE, Scopus == TRUE) |> pull(n_cr) |> sum()` articles were exclusive to both Scopus and hoaddata. Exclusively in hoaddata were `r upset_data |> filter(Crossref == TRUE, WoS == FALSE, Scopus == FALSE) |> pull(n_cr) |> sum()` articles, and exclusively in the proprietary databases were `r upset_data |> filter(Crossref == FALSE, Scopus == TRUE) |> pull(overlap) |> sum(na.rm = TRUE)` articles.

Figure \ref{fig-upset_coverage_results}B also shows DOIs that were only available via hoaddata (grey area).
After validation at the DOI level using the KB databases and manual inspection, the main reasons for missing DOI coverage in the proprietary database were insufficient classification of journal content as original research articles and reviews in hoaddata. Particulalry, Letters and Editorials could not be fully detected.
Paratext recognition also failed in some cases to identify non-scientific content such as journal covers or reviewer lists, which are generally not covered by Scopus and Web of Science. 
To a lesser extent, differences in publication and indexing dates were a reason for non-overlapping DOIs.

@@

### 2025-02-19 analysis/manuscript.tex
Check missing dois

hoaddata, developed and maintained by the author to support open access
monitoring and research (Jahn, 2025), is an R data package comprising
information about the uptake of hybrid open access since 2017 collected
from several openly available data sources. It combines article-level
metadata from Crossref and OpenAlex with transformative agreement
information from the cOAlition S Journal Checker Tool (JCT), which links
journal and institutional data to agreements in the ESAC registry.

@@

Article metadata was retrieved from each databse using all ISSN variants
linked to an ISSN-L. The metadata included DOIs, publication dates,
article types, open access information as well as author roles and
institutional affiliations. Publication years were determined using the
earliest known date of publication in a journal. articles in hoaddata,
this corresponded to Crossref's issued date. For Web of Science and
Scopus, the earliest publication date was used where available, with
Scopus dates specifically determined by the KB through version tracking
of the raw data.

@@

While hoaddata only covers articles with a DOI, Scopus and Web of
Science publication metrics were aggregated using the database
identifier. A subsequent comparison of DOI coverage shows that non-core
articles in Web of Science often lacked a DOI. This was particularly the
case for meeting abstracts, which are notably prevalent in Health
Sciences journals (Melero-Fuentes et al., 2025). Open access indicators
were aggregated by DOI, as Unpaywall only collects information on open
access status for articles with a DOI. A closer look at the core
articles with affiliation data reveals a lack of data on corresponding
authors in the case of OpenAlex, the affiliation data source used by
hoaddata, compared to Web of Science and Scopus. Only about two-thirds
of the articles examined provided corresponding author affiliations. For
first authors, the proportion was 89\%. Therefore, only first author
data for hoaddata were considered in the following analysis.

@@

Journal coverage analysis revealed that 66\% (n = 7,970) of hybrid
journals included in transformative agreements were indexed in all three
databases ( \ref{fig-upset_coverage_results}A). The second-largest set
consisted of journals indexed in both hoaddata and Scopus, comprising
21\% (n = 2,595) of hybrid journals. Notably, 6\% (n = 739) of journals
were exclusively contained in hoaadata, while 3\% (n = 354) were only
found in the proprietory databases Scopus and Web of Science. Upon
insepction, this latter group represents hybrid journals for which no
open access evidence could be retrieved from Crossref, which serves as
the open access evidence source for hoaddata.

In terms of article coverage, Fig B shows the DOI overlap,
distinguihsing between a shared corpus and just those where a DOI was
only available through hoaddata. Upon inspection, reasons for lackign
DOI coverage in proprietory database were different publication dates
and differenc ein document type definitions. As such, demosntrates
certain uncertainies using the open approach. Overall, The majority of
articles can be found

@@

>>>>

hoaddata, developed and maintained by the author to support open access
monitoring and research (Jahn, 2025), is an R data package collecting
information on hybrid open access uptake from multiple openly available
data sources. It combines article-level metadata from Crossref and
OpenAlex with transformative agreement information from the cOAlition S
Journal Checker Tool (JCT), which links journal and institutional data
to agreements in the ESAC registry.

@@

Article metadata were retrieved from each database using all ISSN
variants linked to an ISSN-L. The metadata included DOIs, publication
dates, article types, open access information as well as author roles
and institutional affiliations. Publication years were determined using
the earliest known date of publication in a journal. articles in
hoaddata, this corresponded to Crossref's issued date. For Web of
Science and Scopus, the earliest publication date was used where
available, with Scopus dates specifically determined by the KB through
version tracking of the raw data.

@@

While hoaddata only covers articles with a DOI, Scopus and Web of
Science publication metrics were aggregated using the database
identifier. A subsequent comparison of DOI coverage shows that non-core
articles in Web of Science often lacked a DOI. This was particularly the
case for meeting abstracts, which are notably prevalent in Health
Sciences journals (Melero-Fuentes et al., 2025) and are not indexed by
Scopus (Donner, 2017). Open access indicators were aggregated by DOI, as
Unpaywall only collects information on open access status for articles
with a DOI. A closer look at the core articles with affiliation data
reveals a lack of data on corresponding authors in the case of OpenAlex,
the affiliation data source used by hoaddata, compared to Web of Science
and Scopus. Only about two-thirds of the articles examined provided
corresponding author affiliations. For first authors, the proportion was
89\%. Therefore, only first author data for hoaddata were considered in
the following analysis.

@@

Journal coverage analysis revealed that 66\% (n = 7,970) of hybrid
journals included in transformative agreements were indexed in all three
databases (Figure \ref{fig-upset_coverage_results}A). The second-largest
set consisted of journals indexed in both hoaddata and Scopus,
comprising 21\% (n = 2,595) of hybrid journals. Notably, 6\% (n = 739)
of journals were exclusively contained in hoaadata, while 6\% (n = 748)
were only found in the proprietory database Scopus. Of these, 354 were
also available in the Web of Science. Upon inspection, this group of
hybrid journals included only in the proprietary data sources mainly
represents hybrid journals for which no open access evidence could be
retrieved from Crossref, which served as the open access evidence source
for hoaddata.

In terms of article coverage, Figure \ref{fig-upset_coverage_results}B
shows the total publication volume per combination in terms of DOI
availability. The largest set of hybrid journals, which includes all
three data sources, also contains the largest number of articles. In
total, these journals recorded 6,289,687 overlapping articles,
represented by the blue bar. They represent 94\% of core articles with
DOI indexed in the Web of Science sample, and 86\% in Scopus. Another
657,697 articles are exclusive to both Scopus and hoaddata. Exclusively
in hoaddata were 177,110 articles, and exclusively in the proprietary
databases were 325,194 articles.

Figure \ref{fig-upset_coverage_results}B also shows DOIs that were only
available via hoaddata (grey area). After validation at the DOI level
using the KB databases and manual inspection, the main reasons for
missing DOI coverage in the proprietary database were insufficient
classification of journal content as original research articles and
reviews in hoaddata. Paratext recognition also failed in some cases to
identify non-scientific content such as journal covers or reviewer
lists, which are generally not covered by Scopus and Web of Science. To
a lesser extent, differences in publication and indexing dates were a
reason for non-overlapping DOIs.

@@

\bibitem[\citeproctext]{ref-Donner_2017}
Donner, P. (2017). Document type assignment accuracy in the journal
citation index data of {Web of Science}. \emph{Scientometrics},
\emph{113}(1), 219--236. \url{https://doi.org/10.1007/s11192-017-2483-y}
